\documentclass[letterpaper, 11pt]{extarticle}

\input{preamble}
\input{format}
\input{commands}

\begin{document}

\begin{Large}
    \textsf{\textbf{CS361 - HW2}}
\end{Large}

\vspace{1ex}

\textsf{\textbf{Student:}} \text{Shrey Poshiya}\\

Github Link:

\vspace{2ex}


% Problem 1 - Towers of Hanoi
\begin{problem}{}{problem-1}
    In the classic problem of the \textbf{Towers of Hanoi}, you have 3 towers/rods and $n$ disks of different sizes which can slide onto any tower. The puzzle starts with disks sorted in ascending order of size from top to bottom (i.e., each disk sits on top of an even larger one). You have the following constraints:
    \begin{enumerate}
        \item Only one disk can be moved at a time.
        \item A disk is slid off the top of one tower onto another tower.
        \item A disk cannot be placed on top of a smaller disk.
    \end{enumerate}
    Write a recursive solution that prints every individual move in the format: \texttt{"Move disk X from Tower A to Tower C"}, and analyze the time complexity of your algorithm.
\end{problem}



% Problem 2 - Hybrid Sort
\begin{problem}{}{problem-2}
    You've learned that merge sort has $O(n \log n)$ time complexity, but it has overhead from recursive calls. For small arrays, simpler algorithms like insertion sort can actually be faster.

    \vspace{1ex}
    \noindent\textbf{Part A --- Implementation:}

    Implement a hybrid sort that uses merge sort for large subarrays but switches to insertion sort when the subarray size falls below a threshold $k$.

    \vspace{1ex}
    \noindent\textbf{Part B --- Experimentation and Analysis:}
    \begin{enumerate}
        \item Choose at least 3 different array sizes of your own between 100 and 10,000 (for example: 500, 2000, 8000). Pick sizes that run comfortably on your laptop.
        \item For each array size, run your hybrid sort with $k$ values of $1, 2, 4, 8, 16, 32, 64$.
        \item For each combination of $(n, k)$, repeat the sort 5 times on freshly generated random arrays and record the average runtime in milliseconds.
        \item Plot your results with:
        \begin{itemize}
            \item X-axis: threshold value $k$
            \item Y-axis: average runtime in milliseconds
            \item One curve per array size on the same plot
        \end{itemize}
        \item Identify the optimal value of $k$ (lowest runtime) for each array size from your plot.
        \item Explain why an optimal threshold exists, specifically, why both very small and very large values of $k$ lead to worse performance.
    \end{enumerate}
\end{problem}



% Problem 3 - Counting Inversions
\begin{problem}{}{problem-3}
    An \textbf{inversion} in an array is a pair of indices $(i, j)$ where $i < j$ but $\texttt{arr}[i] > \texttt{arr}[j]$. This measures ``how unsorted'' an array is.

    \vspace{1ex}
    \noindent\textbf{Task:} Modify merge sort to count the total number of inversions in $O(n \log n)$ time.

    \vspace{1ex}
    \noindent\textbf{Examples:}
    \begin{itemize}
        \item $[1, 2, 3, 4]$ has 0 inversions (already sorted)
        \item $[4, 3, 2, 1]$ has 6 inversions (reverse sorted)
        \item $[2, 4, 1, 3]$ has 3 inversions: $(\texttt{arr}[0], \texttt{arr}[2]) = (2,1)$, $(\texttt{arr}[1], \texttt{arr}[2]) = (4,1)$, $(\texttt{arr}[1], \texttt{arr}[3]) = (4,3)$, where each pair shows indices $(i,j)$ with $i < j$ and $\texttt{arr}[i] > \texttt{arr}[j]$
    \end{itemize}

    \vspace{1ex}
    \noindent\textbf{Questions:}
    \begin{enumerate}
        \item What is the maximum number of inversions possible for an array of size $n$?
        \item How does the inversion count relate to sorting algorithms like bubble sort?
    \end{enumerate}
\end{problem}



% Problem 4 - Queue Using Two Stacks
\begin{problem}{}{problem-4}
    Show how to implement a queue using two stacks.
\end{problem}



% Problem 5 - Merge Two Sorted Linked Lists
\begin{problem}{}{problem-5}
    You are given the heads of two sorted linked lists \texttt{list1} and \texttt{list2}. Write a solution that merges the two lists into one sorted list. The list should be made by splicing together the nodes of the first two lists. Your function should return the head of the merged linked list.

    \vspace{1ex}
    \noindent\textbf{Example:}\\
    Input: \texttt{list1 = [1,2,4]}, \texttt{list2 = [1,3,4]}\\
    Output: \texttt{[1,1,2,3,4,4]}
\end{problem}



% Problem 6 - Sort Matrix Diagonals
\begin{problem}{}{problem-6}
    A \textbf{matrix diagonal} is a diagonal line of cells starting from some cell in either the topmost row or leftmost column and going in the bottom-right direction until reaching the matrix's end. Write a solution for a given $m \times n$ matrix of integers, sort each matrix diagonal in ascending order and return the resulting matrix.

    \vspace{1ex}
    \noindent\textbf{Example:}

    \[
    \text{Input:}
    \quad
    \begin{pmatrix}
    3 & 3 & 1 & 1 \\
    2 & 2 & 1 & 2 \\
    1 & 1 & 1 & 2
    \end{pmatrix}
    \qquad
    \text{Output:}
    \quad
    \begin{pmatrix}
    1 & 1 & 1 & 1 \\
    1 & 2 & 2 & 2 \\
    1 & 2 & 3 & 3
    \end{pmatrix}
    \]
\end{problem}



% Problem 7 - Binary Search Tree
\begin{problem}{}{problem-7}
    Consider the following sequence of numbers: $15, 10, 20, 8, 12, 17, 25$.

    \begin{enumerate}[(a)]
        \item Insert these numbers in the given order into a Binary Search Tree (BST). Draw the resulting tree.

        \item Identify whether the constructed BST is also a \textbf{Complete Binary Tree} and a \textbf{Full Binary Tree}. Justify your answer.

        \item What is the in-order traversal of the BST you constructed? What do you notice about the output?
    \end{enumerate}
\end{problem}



\end{document}
